Crunching Large Graphs with Commodity Processors

### 2011-01-26 doc/hotpar2011/softxmt-hotpar2011.tex
intro fixes?

The most interesting computational challenge comes from large,
low-diameter, power law graphs: this combination makes extracting
performance difficult. They do not fit in a single
commodity machine's memory. Their low diameter means that they are
difficult to lay out with locality: every vertex needs to be near every
other vertex. Their power law distribution means that a few vertices
generate much more work than the rest, leading to load
imbalances. Together, these properties lead to a
difficult conclusion: each edge traversal is likely to induce a very
expensive cache miss or remote access.

@@

We want the best of both worlds. Our goal is to build a system that
has good performance on these low-locality graph codes but is
implemented with cheap commodity processors.

Figure~\ref{fig:system-overview} shows the structure of our
proposal. It is composed of multiple nodes built with commodity
processors, communicating over an Infiniband network. We add two
components: a runtime system, responsible for executing the user's
threads, and a global memory manager, responsible for facilitating
memory requests to the non-coherent global memory shared across the
nodes.  We are exploring a mix of hardware and software
implementations of the memory manager.

This paper explores the feasibility of our multi-node goal using a
single-node implementation of our lightweight threading runtime. We
use prefetch instructions together with lightweight threads to
tolerate the latency of DRAM access on the local node node. We use
pointer chasing benchmarks to verify our runtime's overhead is
acceptable. We also model the effects of pointer chasing on a remote
node.

The rest of the paper is organized as follows. We describe our
proposal from a programmer's viewpoint in section~\ref{sec:model}. We
describe the implementation of our latency-tolerant runtime and plans
for extending to multiple nodes in section~\ref{sec:approach}. We
evaluate our runtime in section~\ref{sec:evaluation}.  We present
related work in section~\ref{sec:related} and conclude in
section~\ref{sec:conclusion}.

>>>>

The most interesting computational challenge comes from large,
low-diameter, power law graphs: this combination makes extracting
performance difficult. They do not fit in a single
commodity machine's memory. Their low diameter means that they are
difficult to lay out with locality: every vertex needs to be near every
other vertex. Their power law distribution means that a few vertices
generate much more work than the rest, leading to load
imbalances. Together, these properties lead to a
difficult conclusion: each edge traversal is likely to lead to a very
expensive inter-node communication.

@@

We believe we can build a system based mostly on commodity parts that
can attain XMT-like performance with a familiar, also XMT-like,
programming model, but at a fraction of the cost. Therefore, our goal
is to build a system that has good performance on low-locality graph
codes but is implemented with cheap commodity processors and simple
additional support in an FPGA. This approach also has the added
benefit of not sacrificing general-purpose performance.

Figure~\ref{fig:system-overview} shows the structure of our
proposal. It is composed of multiple nodes built with commodity
processors, communicating over an Infiniband network. We add two
components: a runtime system, responsible for executing the user's
threads, and a global memory manager, responsible for facilitating
memory requests to the non-coherent global memory shared across the
nodes.  We are exploring a mix of hardware (FPGA) and software
implementations of the memory manager.

This paper describes the vision of our system and explores the
feasibility of our ideas using a single-node implementation of a
lightweight threading runtime. We use prefetch instructions together
with lightweight threads to tolerate the latency of DRAM access on the
local node node. We use adversarial pointer chasing benchmarks to
verify our runtime's overhead is acceptable. We also model the effects
of pointer chasing on a remote node by artificially increasing
latenciess to the order of several micro-seconds.

The rest of the paper is organized as follows. We quickly describe our
programming model in section~\ref{sec:model}. We describe the
implementation of our latency-tolerant runtime and plans for extending
to multiple nodes in section~\ref{sec:approach}. We evaluate our
runtime in section~\ref{sec:evaluation}.  We present related work in
section~\ref{sec:related} and conclude in
section~\ref{sec:conclusion}.
